---
title: "Data-Related Jobs Salary Analysis"
author: "William Lin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
rm(list=ls()) # Clean environment
```

```{r}
library(ggplot2)
```

Key performance indicator (KPI): Paid Wage Per Year

Team ANOVA will be analyzing the salary dataset for our DS311 project. The salary dataset displays records from 167,278 Labor Condition Applications (which include H1-B and other specialty visas) and the permanent resident applications from 2008 to 2015. The dataset intends to track who is entering and leaving the US government for work-related reasons. Each applicant provides information based on their work location, paid/prevailing wages, and their selected job position. Within this dataset, there are 25 variables with 2 extra non-meaningful variables. Several goals of the project include comparing prevailing wages and paid wages, comparing wages by job title, comparing wages by location, and more.


## Salary data

```{r}
salary <- read.csv("/Users/linw/Desktop/Team-Anova-Project/Will_WIP/salary_data_states.csv")
```

```{r}
# Inspecting the data
dim(salary)
head(salary)
tail(salary)
```

The dataset has 167,278 observations and 27 columns. Two variables are extras or non-meaningful: the initials/codes of U.S. states/territories, and the variable order seems to be an index from a previous uncleaned version of the dataset.

## Summary statistics

```{r}
summary(salary$PREVAILING_WAGE_PER_YEAR)
summary(salary$PAID_WAGE_PER_YEAR)
```

The minimum, first quartile, median, third quartile, and maximum prevailing wage per year are as follows: \$10,504, \$56,880, \$70,928, \$90,459, and \$320,000. The mean prevailing wage per year is \$74274.86 with standard deviation \$25356.24. The numerical summary for the paid wage per year is as follows: \$10,500, \$63,000, \$78,600, \$100,006, and \$2,500,000. The mean paid wage per year is \$85,532.77 with standard deviation $38738.47. 

## Data cleaning

```{r}
# Before cleaning
boxplot(salary$PAID_WAGE_PER_YEAR)
hist(salary$PAID_WAGE_PER_YEAR, 
     main='Histogram of Paid Wage Per Year', 
     xlab='Paid Wage Per Year')

#
```


## How do offered salaries compare to the prevailing wage?

We create a new feature "PAY_DIFFERENCE", which represents the difference between the paid wage per year and the prevailing wage per year.

```{r}
salary$PAY_DIFFERENCE = salary$PAID_WAGE_PER_YEAR - salary$PREVAILING_WAGE_PER_YEAR

# Summary stats: Difference in wage (paid vs. prevailing)
summary(salary$PAY_DIFFERENCE)
```

```{r}
# Boxplot in base R
boxplot(salary$PAY_DIFFERENCE)

# Create box plot with ggplot2
ggplot(data=salary, aes(y=PAY_DIFFERENCE, x=1)) + geom_boxplot()

# Histogram with ggplot2
ggplot(data=salary) + geom_histogram(aes(x=PAY_DIFFERENCE))
```

- Are there job sub-categories that tend to get overpaid or under-paid?

```{r}
ggplot(data=salary, aes(x=JOB_TITLE_SUBGROUP,
                        y=PAY_DIFFERENCE)) + 
  geom_boxplot() +
  coord_flip()

# Paid Wage by job sub-title
ggplot(data=salary, aes(x=JOB_TITLE_SUBGROUP,
                        y=PAID_WAGE_PER_YEAR)) + 
  geom_boxplot() +
  coord_flip() + 
  ylim(-50000, 200000)

# Difference between paid and prevailing wage
ggplot(data=salary, aes(x=JOB_TITLE_SUBGROUP,
                        y=PAY_DIFFERENCE)) + 
  geom_boxplot() +
  coord_flip() + 
  ylim(-75000, 200000)

# Subset 
bus_analyst_subset = subset(salary, JOB_TITLE_SUBGROUP == "business analyst")
summary(bus_analyst_subset$PAID_WAGE_PER_YEAR)
```


- Are there companies that tend to over-pay or under-pay?

```{r}
faang_subset = subset(salary, 
                      EMPLOYER_NAME == "FACEBOOK, INC." | 
                        EMPLOYER_NAME == "APPLE INC." | 
                        EMPLOYER_NAME == "AMAZON CORPORATE LLC" | 
                        EMPLOYER_NAME == "NETFLIX, INC." | 
                        EMPLOYER_NAME == "GOOGLE, INC.")

# Difference between paid and prevailing wage, by employer
ggplot(data=faang_subset, aes(x=EMPLOYER_NAME,
                        y=PAY_DIFFERENCE)) + 
  geom_boxplot() +
  coord_flip()
```

-  Will the answer change if I take standard of living into account?

```{r}

```

## Paid Wage by Required Education Level

```{r}
# Create factor variable: equation level required (7 levels)
education = as.factor(salary$EDUCATION_LEVEL_REQUIRED)
education = relevel(factor(education, levels=c("None", "High School", 
                                               "Associate's", "Bachelor's", 
                                               "Master's", "Doctorate", 
                                               "Other")), 
                                      ref = "None")

# Plot
ggplot(data=salary, aes(x=education,
                        y=PAID_WAGE_PER_YEAR)) + 
  geom_boxplot() +
  coord_flip() +
  ylim(-20000, 300000) + 
  xlab("Required Education Level") + 
  ylab("Paid Wage Per Year")

# Ascending Order
ggplot(data=salary, aes(x=reorder(education, PAID_WAGE_PER_YEAR, FUN = median),
                        y=PAID_WAGE_PER_YEAR)) + 
  geom_boxplot() +
  coord_flip() +
  ylim(-20000, 300000) + 
  xlab("Required Education Level") + 
  ylab("Paid Wage Per Year")
```


## How is wage correlated with experience required?

```{r}
EXPERIENCE_REQUIRED_YEARS = salary$EXPERIENCE_REQUIRED_NUM_MONTHS / 12

# Scatter plot with ggplot2
ggplot(data=salary, aes(x=EXPERIENCE_REQUIRED_YEARS, 
                        y=PAID_WAGE_PER_YEAR)) + 
  geom_point() + 
  ylab("Paid Wage Per Year") + 
  xlab("Experience Required (Years)") + 
  ylim(0, 300000) + 
  geom_smooth(method=lm)
```
