---
title: "Salary Data Analysis"
author: "Tristan Lampley"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE}
rm(list=ls()) # Clean environment
```

```{r}
library(ggplot2)
```

## Salary data

```{r}
salary <- read.csv("C:/Users/kurox/Desktop/DS311/Team_Anova/Team-Anova-Project/Tristan_WIP/salary_data_states.csv")
```

```{r}
# Inspecting the data
dim(salary)
head(salary)
tail(salary)
```

## Summary statistics

```{r}
summary(salary$PREVAILING_WAGE_PER_YEAR)
summary(salary$PAID_WAGE_PER_YEAR)
```

The variables we are mainly concerned with are prevailing wage per year and paid wage per year. The minimum, first quartile, median, third quartile, and maximum prevailing wage per year are as follows: \$10,504, \$56,880, \$70,928, \$90,459, and \$320,000. The mean prevailing wage per year is \$74274.86 with standard deviation \$25356.24. The numerical summary for the paid wage per year is as follows: \$10,500, \$63,000, \$78,600, \$100,006, and \$2,500,000. The mean paid wage per year is \$85,532.77 with standard deviation $38738.47. 

## Data cleaning

```{r}
# Before cleaning
boxplot(salary$)


```

## How do offered salaries compare to the prevailing wage?

We create new feature "PAY_DIFFERENCE", which is representing the difference between paid wage per year and prevailing wage per year.

```{r}
# salary$PAY_DIFFERENCE = salary$Paid_WAGE_PER_YEAR - salary$PREVAILING_WAGE_PER_YEAR

summary(salary$PAY_DIFFERENCE)

#Creating boxplot

boxplot(salary$PAY_DIFFERENCE)



```

- Are there job sub-categories that tend to get overpaid or under-paid?

```{r}

```

- Are there companies that tend to over-pay or under-pay?

```{r}

```

-  Will the answer change if I take standard of living into account?

```{r}
```

## Paid Wage by College Degree

```{r}

```


